---
title: "Project flowers"
format: html
execute: 
  warning: false
  echo: false
---

```{r}
library(tidyverse)

lc.gf <- read.csv("data/lcareas_gf2021.csv")
lc.hw <- read.csv("data/lcareas_hw2025.csv")
lc.iv <- read.csv("data/lcareas_iverson.csv")

flr.gf <- read.csv("data/florlc_gf2021.csv")
flr.hw <- read.csv("data/florlc_hw2025.csv")

lcflr.gf <- read.csv("data/flowerdens_gf2021.csv")
lcflr.iv <- read.csv("data/flowerdens_iverson.csv")
lcflr.hw <- read.csv("data/flowerdens_hw2025.csv")

key.iv <- read.csv("data-raw/Iversonetal/iverson_landcover_codes.csv")
lckey.hw <- read.csv("data/spatial/hw2025_lc_key.csv")

```

## Hemberger & Williams 2025

```{r}

ggplot(lcflr.hw, aes(month, flowerdens, color = raster.name)) +
  geom_point() + geom_line()

```

### Compare landscape buffers

Panels represent quartiles of percent habitat within 750 m buffer.

```{r}

buf.hw <- lc.hw |> filter(buffer == 750)

cropbuf.hw <- buf.hw |>
  pivot_longer(cols = CROPLA:RICE, names_to = "raster.name", values_to = "area") |>
  left_join(lcflr.hw, by = "raster.name", relationship = "many-to-many") |>
  filter(!is.na(period) & !is.na(area)) |>
  mutate(total.flowers = area*flowerdens)

croparea.hw <- cropbuf.hw |>
  group_by(period, month, SITE, landcover.4cat) |>
  summarize(area = sum(area), total.flowers = sum(total.flowers), .groups = "drop") |>
  pivot_wider(id_cols = c(period, month, SITE), 
              names_from = landcover.4cat, values_from = c(area, total.flowers)) |>
  mutate(
    total.flowers = total.flowers_crop + total.flowers_natural,
    total.area = area_crop + area_natural,
    pct.natural = 100*area_natural/(750*750*pi),
    pct.nat.intv = cut_width(pct.natural, width = 25, boundary = 0)
  )

croparea.hw.mn <- croparea.hw |> group_by(month, pct.nat.intv) |>
  summarize(total.flowers = mean(total.flowers), .groups = "drop")

ggplot(croparea.hw, aes(month, total.flowers)) +
  geom_point(aes(group = SITE, color = pct.natural)) + geom_line(aes(group = SITE, color = pct.natural)) + 
  geom_line(data = croparea.hw.mn, color = "red", lwd = 1, lty = 5) +
  scale_color_viridis_c(option = "viridis") + facet_wrap(~pct.nat.intv)

```

### Crop richness

```{r}

croprich.hw <- cropbuf.hw |> rowwise() |>
  mutate(crop = landcover.4cat=="crop" & area > 0) |>
  group_by(period, month, SITE) |>
  summarize(total.flowers = sum(total.flowers), crop.rich = sum(crop), 
            .groups = "drop")

croparearich.hw <- croparea.hw |> select(month, period, SITE, area_crop, area_natural, pct.natural, total.flowers) |>
  left_join(croprich.hw |> select(month, period, SITE, crop.rich), by = c("month", "period", "SITE"))

croparearich.hw.mn <- croparearich.hw |> group_by(month, crop.rich) |>
  summarize(total.flowers = mean(total.flowers), .groups = "drop")

ggplot(croparearich.hw, aes(month, total.flowers)) + 
  geom_point(aes(group = SITE, color = pct.natural)) + geom_line(aes(group = SITE, color = pct.natural)) + 
  geom_line(data = croparearich.hw.mn, color = "red", lwd = 1, lty = 5) +
  facet_wrap(~crop.rich) + scale_color_viridis_c(option = "viridis")
```

### Crop diversity

```{r}

cropdiv.hw <- croparea.hw |> select(SITE, area_crop) |> distinct() |>
  right_join(cropbuf.hw, by = "SITE") |> filter(landcover.4cat == "crop") |> 
  mutate(P_i = area/area_crop, P_i2 = P_i * log(P_i)) |>
  group_by(period, month, SITE) |>
  summarize(crop.shandiv = - sum(P_i2, na.rm = TRUE), .groups = "drop") |>
  right_join(croparearich.hw, by = c("period", "month", "SITE"))

ggplot(cropdiv.hw, aes(month, total.flowers)) + 
  geom_point(aes(group = SITE, color = crop.shandiv)) + geom_line(aes(group = SITE, color = crop.shandiv)) +
  geom_line(data = croparearich.hw.mn, color = "red", lwd = 1, lty = 5) + facet_wrap(~crop.rich) +
  scale_color_viridis_c(option = "viridis")

```

### crop richness vs % natural correlation

```{r}

ggplot(croparearich.hw, aes(crop.rich, pct.natural)) + geom_point()

```

### crop richness vs crop shannon diversity

```{r}

ggplot(cropdiv.hw, aes(crop.rich, crop.shandiv)) + geom_point()

```

## Guezen and Forrest 2021

```{r}

ggplot(lcflr.gf,
       aes(datePeriod, flowerdens, color = landtype)) + geom_point() + geom_line()

```

### Compare landscape buffers

Panels represent quartiles of percent habitat within 750 m buffer.

```{r}

buf.gf <- lc.gf |> filter(buffer == 750)

cropbuf.gf <- buf.gf |>
  pivot_longer(cols = Blackberry:Canola, names_to = "landtype", values_to = "area") |> 
  mutate(landtype = gsub("\\.","-", tolower(landtype))) |>
  left_join(lcflr.gf, by = "landtype", relationship = "many-to-many") |>
  filter(!is.na(datePeriod) & !is.na(area)) |>
  mutate(total.flowers = area*flowerdens)

croparea.gf <- cropbuf.gf |>
  group_by(datePeriod, transect, crop) |>
  summarize(area = sum(area), total.flowers = sum(total.flowers), .groups = "drop") |>
  pivot_wider(id_cols = c(datePeriod, transect), 
              names_from = crop, values_from = c(area, total.flowers)) |>
  mutate(
    total.flowers = total.flowers_crop + total.flowers_natural,
    total.area = area_crop + area_natural,
    pct.natural = 100*area_natural/(750*750*pi),
    pct.nat.intv = cut_width(pct.natural, width = 25, boundary = 0)
  )

croparea.gf.mn <- croparea.gf |> group_by(datePeriod, pct.nat.intv) |>
  summarize(total.flowers = mean(total.flowers), .groups = "drop")

ggplot(croparea.gf, aes(datePeriod, total.flowers)) +
  geom_point(aes(group = transect, color = pct.natural)) + geom_line(aes(group = transect, color = pct.natural)) + 
  geom_line(data = croparea.gf.mn, color = "red", lwd = 1, lty = 5) +
  scale_color_viridis_c(option = "viridis") + facet_wrap(~pct.nat.intv)


```

### Crop richness

```{r}

croprich.gf <- cropbuf.gf |> rowwise() |>
  mutate(crop = crop=="crop" & area > 0) |>
  group_by(datePeriod, transect) |>
  summarize(total.flowers = sum(total.flowers), crop.rich = sum(crop), 
            .groups = "drop")

croparearich.gf <- croparea.gf |> select(datePeriod, transect, area_crop, area_natural, pct.natural, total.flowers) |>
  left_join(croprich.gf |> select(datePeriod, transect, crop.rich), by = c("datePeriod", "transect"))

croparearich.gf.mn <- croparearich.gf |> group_by(datePeriod, crop.rich) |>
  summarize(total.flowers = mean(total.flowers), .groups = "drop")

ggplot(croparearich.gf, aes(datePeriod, total.flowers)) + 
  geom_point(aes(group = transect, color = pct.natural)) + geom_line(aes(group = transect, color = pct.natural)) + 
  geom_line(data = croparearich.gf.mn, color = "red", lwd = 1, lty = 5) +
  facet_wrap(~crop.rich) + scale_color_viridis_c(option = "viridis")

```

### Crop diversity

```{r}

cropdiv.gf <- croparea.gf |> select(transect, area_crop) |> distinct() |>
  right_join(cropbuf.gf, by = "transect") |> filter(crop == "crop") |> 
  mutate(P_i = area/area_crop, P_i2 = P_i * log(P_i)) |>
  group_by(datePeriod, transect) |>
  summarize(crop.shandiv = - sum(P_i2, na.rm = TRUE), .groups = "drop") |>
  right_join(croparearich.gf, by = c("datePeriod", "transect"))

ggplot(cropdiv.gf, aes(datePeriod, total.flowers)) + 
  geom_point(aes(group = transect, color = crop.shandiv)) + geom_line(aes(group = transect, color = crop.shandiv)) + 
  geom_line(data = croparearich.gf.mn, color = "red", lwd = 1, lty = 5) +
  facet_wrap(~crop.rich) + scale_color_viridis_c(option = "viridis")

```

### crop richness vs % natural correlation

Guezen & Forrest points are transects clustered by sites. At 750 m, some buffers are overlapping, some are not.

```{r}

ggplot(croparearich.gf |> 
         separate_wider_delim(transect, delim = ".", names = c("site"), too_many = "drop", cols_remove = FALSE),
       aes(crop.rich, pct.natural, color = site)) + geom_point()

```

### crop richness vs crop shannon diversity

Guezen & Forrest points are transects clustered by sites. At 750 m, some buffers are overlapping, some are not.

```{r}

ggplot(cropdiv.gf |> 
         separate_wider_delim(transect, delim = ".", names = c("site"), too_many = "drop", cols_remove = FALSE),
       aes(crop.rich, crop.shandiv, color = site)) + geom_point()

```

## Iverson data

Aaron took surveys of all vegetation within a larger area using a modified Whitaker plot. Herbaceous plants were measured by coverage and trees were counted. Flowers were not counted during the survey but instead counted at peak flowering density independently of the site. The average peak flower density (or per-individual count for trees) was used to define the flowering curve that was calculated for the site over the entire year. Aaron's workflow calculates total flower area per ha using floral dimensions from literature or field measurements, but I stop at flower counts to match the other studies.

### Floral density curves of individual land covers

```{r}
#| fig-height: 6
#| fig-width: 8.5

ggplot(lcflr.iv |> 
         pivot_longer(d1:d365, names_to = "julianDay", values_to = "flowers_per_m2") |>
         mutate(julianDay = as.numeric(gsub("d","",julianDay))),
       aes(julianDay, flowers_per_m2,
           group = Floral_resources_landcover, 
           color = Floral_resources_landcover)) + geom_line() +
  theme(legend.position = "bottom", 
        legend.title=element_blank())

```

### Compare landscape buffers

Panels represent quartiles of percent habitat within 750 m buffer.

```{r}

buf.iv <- lc.iv |> filter(buffer == "buf750")

cropbuf.iv <- buf.iv |>
  pivot_longer(cols = Alfalfa:Plums,
               names_to = "Floral_resources_landcover",
               values_to = "area") |> 
  mutate(Floral_resources_landcover = 
           replace_values(Floral_resources_landcover,
                          "Urban.tree" ~ "Urban tree")) |>
  left_join(lcflr.iv, by = "Floral_resources_landcover") |>
  filter(!is.na(area)) |> 
  pivot_longer(d1:d365, names_to = "julianDay", values_to = "flowers_m2") |>
  mutate(
    total.flowers = area * flowers_m2,
    julianDay = as.numeric(gsub("d","",julianDay))
    )


croparea.iv <- cropbuf.iv |>
  group_by(site, Group, julianDay) |>
  summarize(area = sum(area), total.flowers = sum(total.flowers), .groups = "drop") |>
  pivot_wider(id_cols = c(julianDay, site), 
              names_from = Group, values_from = c(area, total.flowers)) |>
  mutate(
    total.flowers = rowSums(across(total.flowers_Agriculture:total.flowers_Wetland)),
    total.area = rowSums(across(area_Agriculture:area_Wetland)),
    area_natural = area_Forest + area_Wetland + area_Succesional,
    pct.natural = 100*area_natural/(750*750*pi),
    pct.nat.intv = cut_width(pct.natural, width = 25, boundary = 0)
  )

croparea.iv.sum <- croparea.iv |> select(site, pct.natural, pct.nat.intv) |> distinct()

croparea.iv.mn <- croparea.iv |> group_by(julianDay, pct.nat.intv) |>
  summarize(total.flowers = mean(total.flowers), .groups = "drop")

ggplot(croparea.iv, aes(julianDay, total.flowers)) +
  geom_line(aes(group = site, color = pct.natural)) + 
  geom_line(data = croparea.iv.mn, mapping = aes(julianDay, total.flowers), color = "red", lwd = 1) +
  scale_color_viridis_c(option = "viridis") + facet_wrap(~pct.nat.intv)


```

### Crop richness

```{r}

croprich.iv <- cropbuf.iv |> rowwise() |>
  mutate(crop = Group=="Agriculture" & area > 0) |>
  group_by(julianDay, site) |>
  summarize(total.flowers = sum(total.flowers), crop.rich = sum(crop), .groups = "drop")

croparearich.iv <- croparea.iv |> select(julianDay, site, area_Agriculture, area_natural, pct.natural, total.flowers) |>
  rename(area_crop = area_Agriculture) |>
  left_join(croprich.iv |> select(julianDay, site, crop.rich), by = c("julianDay", "site"))

croparearich.iv.mn <- croparearich.iv |> group_by(julianDay, crop.rich) |>
  summarize(total.flowers = mean(total.flowers), .groups = "drop")

ggplot(croparearich.iv, aes(julianDay, total.flowers)) + 
  geom_line(aes(group = site, color = pct.natural)) +
  geom_line(data = croparearich.iv.mn, mapping = aes(julianDay, total.flowers), color = "red", lwd = 0.5, lty = 5) +
  facet_wrap(~crop.rich) + scale_color_viridis_c(option = "viridis")

```

### Crop diversity

```{r}

cropdiv.iv <- croparea.iv |> select(site, area_Agriculture) |> distinct() |>
  right_join(cropbuf.iv, by = "site") |> filter(Group == "Agriculture") |> 
  mutate(P_i = area/area_Agriculture, P_i2 = P_i * log(P_i)) |>
  group_by(julianDay, site) |>
  summarize(crop.shandiv = - sum(P_i2, na.rm = TRUE), .groups = "drop") |>
  right_join(croparearich.iv, by = c("julianDay", "site"))

ggplot(cropdiv.iv, aes(julianDay, total.flowers)) + 
  geom_line(aes(group = site, color = crop.shandiv)) +
  geom_line(data = croparearich.iv.mn, mapping = aes(julianDay, total.flowers), color = "red", lwd = 0.5, lty = 5) +
  facet_wrap(~crop.rich) + scale_color_viridis_c(option = "viridis")

```

### crop richness vs % natural correlation

Iverson points are sites clustered by farm. At 750 m, some buffers are overlapping.

Color + shape combinations represent shared farms in these plots.

```{r}

cropdiv.iv.sum <- cropdiv.iv |> select(site, pct.natural, crop.rich, crop.shandiv) |> 
  distinct() |> 
  separate_wider_delim(site, delim = " ", names = c("farmsite", "farm"),
                       too_many = "merge", cols_remove = FALSE)

# Source - https://stackoverflow.com/q/76810554
# Posted by Lucile Mo', modified by community. See post 'Timeline' for change history
# Retrieved 2026-09-09, License - CC BY-SA 4.0

shapes <- c(0,2, 4, 7, 10, 15, 17, 18)
col_vibrant <- c("#0077BB","#33BBEE","#009988","#EE7733","#CC3311","#EE3377",  "#BBBBBB")

ggplot(cropdiv.iv.sum, aes(crop.rich, pct.natural, color = farm, pch = farm)) + geom_point() +
  scale_colour_manual(values = sample(col_vibrant, 81, replace = TRUE)) +
  scale_shape_manual(values = sample(shapes, 81, replace = TRUE)) +
  theme(legend.position = "none")

```

### crop richness vs crop shannon diversity

```{r}

ggplot(cropdiv.iv.sum, aes(crop.rich, crop.shandiv, color = farm, pch = farm)) + geom_point() +
  scale_colour_manual(values = sample(col_vibrant, 81, replace = TRUE)) +
  scale_shape_manual(values = sample(shapes, 81, replace = TRUE)) +
  theme(legend.position = "none")

```

```{r}
#| include: false

write.csv(cropdiv.gf, "data/cropdiv_gf2021.csv", row.names = FALSE)
write.csv(cropdiv.hw, "data/cropdiv_hw2025.csv", row.names = FALSE)
write.csv(cropdiv.iv, "data/cropdiv_iverson.csv", row.names = FALSE)

```
